Solar wind prediction for the Parker Solar Probe orbit

### 2017-06-17 paperMVVB/sections/content.tex
included figure yearly exponents; sect. 7.2 beginning modified; text: some wording changed

The fitting of $W_\text{II}(x)$ to the velocity's frequency distribution gives the values of the now five fit parameters ($c$, $x_\text{med,1}$, $x_\text{avg,1}$, $x_\text{med,2}$ and $x_\text{avg,2}$), which are also listed in Table~\ref{tab:lognormal_fit_parameters} together with the median and mean of the composed distribution, which can be derived via solving
\begin{align}
	\int W_\text{II}(x)\,\text{d}x = 0	&	&\text{and}	&	&\int x\,W_\text{II}(x)\,\text{d}x = 0
\end{align}
respectively.\\
As anticipated, this more complex fit function is more accurate in describing the velocity's frequency distribution (see Fig.~\ref{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}).
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/histogram_fits_V_a_zoom_dbl_paper_pdfplot.pdf}}
	\caption{Plot of the velocity's frequency distribution (same as in Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}) and its compositional lognormal fit. The fit's median and mean values and its two fit parts are indicated as well. The inset only has a zoomed-in frequency axis, its x-axis stays the same.}
	\label{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}
\end{figure}
That is why we keep using the double lognormal ansatz for the velocity frequency fits in the following sections.\\
In this static model the slow and fast part contribute almost equally ($c \approx 0.5$), which of course is only valid for this kind of long-term average. At different times in a solar cycle their contributions vary strongly.

@@

%solar activity = SSN $\neq$ solar cycle\\
The solar wind velocity and its close friends density and temperature are known to depend on the state of the solar cycle \citep{Schwenn1983}, which is why they follow the SSN indirectly (with time lag).
%density modulation with SSN \citep{Schwenn1983} p.~499\\
Thus we derive the correlation coefficients for different time lags between solar wind parameters and SSN (see Fig.~\ref{fig:OMNI_yearly_ssn_correlation_c_plot}).
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/OMNI_yearly_ssn_correlation_c_plot.pdf}}
	\caption{Plot of the solar wind parameter yearly medians from OMNI data and the yearly SSN from the \citet{sidc} with cycle number (top). Their correlation coefficients with the yearly SSN for time lags back to -15 years (bottom).}
	\label{fig:OMNI_yearly_ssn_correlation_c_plot}
\end{figure}

The highest correlation coefficient for the magnetic field strength is 0.728. It is without lag time and the highest of all solar wind parameters, which is expected because the SSN is directly proportional to the magnetic flux \citep{Smith2003}.\\
Velocity and temperature have a lag time of 3~years with their maximal correlation coefficients (0.453 and 0.540). The density has a lag time of 6~years (0.468), which is in agreement with the by \citet{Bougeret1984} documented density anticorrelation with SSN.	%p.~406\\

% maximal correlation coefficients and their lag times:\\
% magnetic field strength: lag 0, 0.728\\
% velocity: lag -3, 0.453\\
% density: lag -6, 0.468\\
% temperature: lag -3, 0.540\\
For all parameters their correlation coefficients' amplitudes decline with increasing lag time and, as also expected, show a frequency of about 11~years.

@@

Naturally, the fit models match with the general data trends, though single year variations are not able to be replicated by the model (e.g. the high velocity and temperature values in 1974, 1994 and 2003) (see Fig.~\ref{fig:OMNI_yearly_BVdblNTSSN_fit_e_plot}).
\begin{figure*}
	\includegraphics[width=18cm]{figures/OMNI_yearly_BVdblNTSSN_fit_e_plot.pdf}
	\caption{Solar wind parameter data frequencies, lognormal fit models with their median values (white) and the corresponding yearly SSN (grey) over the OMNI time period 1963--2016. The for the models shifted SSN is indicated by a black line. The velocity median is derived from the SSN weighted constant lognormal parts (dotted).}
	\label{fig:OMNI_yearly_BVdblNTSSN_fit_e_plot}
\end{figure*}
The comparison with the yearly data median values over SSN shows that the from the model obtained medians have a quite similar slope (see Fig.~\ref{fig:OMNI_yearly_BVNTvsSSN_a}).\\
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/OMNI_yearly_BVNTvsSSN_a.pdf}}
	\caption{Solar wind parameter median over lagged SSN. The yearly data medians (+) with their weighted linear fit (solid) are obtained from OMNI data. The error bars denote the SSN standard deviation and the relative weight from the yearly data coverage. The SSN dependent median (\ref{eq:median_with_ssn}) is from the lognormal model fit (dashed). For the velocity the median is derived from the SSN weighting of the slow and fast model parts, whose magnitudes are SSN independent (dotted).}
	\label{fig:OMNI_yearly_BVNTvsSSN_a}
\end{figure}

Again, the velocity gets a special treatment with the double lognormal distribution (\ref{eq:double_lognormal_fit_function}). It is known that slow and fast solar wind stream occurrence rates follow the solar cycle, yet their magnitudes stay fairly stable (cite?). Thus we keep the two velocity components' positions constant and vary instead their balance with the SSN:
\begin{align}
	c(ssn) &= c_a \cdot ssn + c_b\,.	\label{eq:balance_with_ssn}
\end{align}
The fit result (see Table~\ref{tab:ssn_fit_parameters}) is a model in which three years after solar cycle minimum (SSN of zero) the slow solar wind has a share of almost two-thirds and decreases further with increasing SSN (see Fig.~\ref{fig:Vdbl_SSN_ratio_f_plot}).\\
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/Vdbl_SSN_ratio_f_plot.pdf}}
	\caption{Balance of slow to fast solar wind over the by 3~years lagged SSN. The yearly ratios (+) and their weighted linear fit (solid) are obtained from OMNI data with a threshold velocity of $v_\text{th} = \SI{400}{\km\per\s}$. The error bars denote the SSN standard deviation and the relative weight from the yearly data coverage. The model's balance parameter (\ref{eq:balance_with_ssn}) and derived ratio (same threshold) are plotted as dashed and dotted lines.}
	\label{fig:Vdbl_SSN_ratio_f_plot}
\end{figure}
% 3 years after minimum (SSN of zero) the slow wind has its maximal share of 0.638(32). 3 years after maximum (SSN of 200) it has only a share of 0.278(37).\\
To compare the ratios of slow to fast wind between model and data, we simply apply the commonly used constant velocity threshold of $v_\text{th} = \SI{400}{\km\per\s}$ (cite?). The linear fit to the yearly data ratio and the derived model ratio are quite similar (see Fig.~\ref{fig:Vdbl_SSN_ratio_f_plot}). Specific velocity thresholds between slow and fast solar wind cannot be directly compared with the somewhat steeper balance parameter of this model. The model's balance may represent the actual ratios of the solar wind types in a more realistic way than a specific velocity threshold does, since the velocity ranges of both types overlap \citep{McGregor2011b}.
%(i.e. there exists slow wind with properties of fast wind and vice versa)

@@

%Helios data biased towards solar minimum
Using this data, we also have to keep in mind that its time coverage is unequally distributed over the solar cycle. Dividing the data by the transition from cycle minimum to maximum (mid 1977) and considering the data gap distributions, the Helios data covers about \SI{68}{\percent} during cycle minimum whereas during maximum only \SI{38}{\percent}.\\

@@

For calculating the median and mean values at different solar distances the data is binned into \SI{0.01}{\au} bins, which is also the native precision in this data set.\\

\subsection{Exponential fitting}
\label{sec:exponential_fitting}
%why exponential function?
An exponential distance behavior is expected from all four parameters (cites?). Therefore we use the exponential function
\begin{align}
	x(r) = d\,r^e	\label{eq:exponential_function}
\end{align}
with the solar distance $r$ for the regression fit of the median and mean. The fits are weighted by data counts per bin.
%fit result table and figure
With $r$ in astronomical units we get the fit coefficients ($d_\text{med}$, $d_\text{avg}$, $e_\text{med}$ and $e_\text{avg}$) as given in Table~\ref{tab:mean_median_fit_parameter}.
\begin{table*}
	\caption{Fit coefficients for the median and mean solar distance dependencies of the four parameters from the combined Helios data set. The errors in brackets are the estimated standard deviations of each fit parameter. The crossing distance is the point where the fitted median and mean intersect.}
	\label{tab:mean_median_fit_parameter}
	\centering
	\begin{tabular}{l
	S[table-format = 1.3(2)]
	S[table-format = +1.3(2)]
	@{}c
	S[table-format = 1.3(2)]
	S[table-format = +1.3(2)]
	S[table-format = 1.1(2)e1]}
		\hline\hline
		\multirow{2}{*}{Parameter}	&\multicolumn{2}{c}{Median}	&	&\multicolumn{2}{c}{Mean}	&\multicolumn{1}{c}{Crossing distance}\\
		\cline{2-3}	\cline{5-6}
			&\multicolumn{1}{c}{$d_\text{med}$\tablefootmark{a}}	&\multicolumn{1}{c}{$e_\text{med}$}	&	&\multicolumn{1}{c}{$d_\text{avg}$\tablefootmark{a}}	&\multicolumn{1}{c}{$e_\text{avg}$}	&\multicolumn{1}{c}{[\si{\au}]}\\
		\hline
		Magnetic field	&5.377(92)	&-1.655(17)	&	&6.05(10)	&-1.546(18)	&0.339(11)\\
		Velocity	&4.107(28)	&0.058(13)	&	&4.356(24)	&0.049(10)	&0.7(83)e3\\
		Density		&5.61(27)	&-2.093(46)	&	&7.57(30)	&-2.010(38)	&0.027(73)\\
		Temperature	&7.14(23)	&-0.913(39)	&	&9.67(21)	&-0.792(28)	&0.082(85)\\
		\hline
	\end{tabular}
	\tablefoot{
		\tablefoottext{a}{Values in their respective units \si{nT}, \SI{e2}{\km\per\s}, \si{\per\cm\cubed} and \SI{e4}{\K}.}
	}
\end{table*}

@@

As mentioned before, we set the exponents of median and average to be identical. Implementing the exponential distance dependency~(\ref{eq:exponential_function}) into the lognormal function (\ref{eq:single_lognormal_fit_function}), we get three fit parameters ($d'_\text{med}$, $d'_\text{avg}$ and the common exponent $e'$).

Naturally, we use the double lognormal function~(\ref{eq:double_lognormal_fit_function}) for the velocity distribution fit, resulting in $W''_\text{II}(x,r)$. The additional fit parameters are the balancing parameter $c'$ and from the second lognormal part $d'_\text{med,2}$ and $d'_\text{avg,2}$. The resulting fit coefficients for the four solar wind parameters are presented in Table~\ref{tab:extrapolation_model_fit_parameters}.
\begin{table*}
	\caption{Fit coefficients from the single lognormal exponential function, respectively double lognormal for the velocity (combined Helios data). The errors in brackets are the estimated standard deviations of each fit parameter.}
	\label{tab:extrapolation_model_fit_parameters}
	\centering
	\sisetup{table-figures-integer=1, table-figures-decimal=3, table-figures-exponent=0}
	\begin{tabular}{l c
	S[table-format = 1.3(2), table-space-text-post = a, table-align-text-post = false]
	S[table-format = 1.3(2), table-space-text-post = a, table-align-text-post = false]
	S[table-format = +1.4(2)]
	S[table-format = 1.3(2)]}
		\hline\hline
		\multicolumn{2}{l}{\multirow{2}{*}{Parameter}}	&\multicolumn{1}{c}{Median\tablefootmark{a}}	&\multicolumn{1}{c}{Mean\tablefootmark{a}}	&\multicolumn{1}{c}{Exponent}	&\multicolumn{1}{c}{Balance}\\
			&	&\multicolumn{1}{c}{$d'_\text{med}$}	&\multicolumn{1}{c}{$d'_\text{avg}$}	&\multicolumn{1}{c}{$e'$}	&$c'$\\
		\hline
		\multicolumn{2}{l}{Magnetic field}	&5.358(25)	&5.705(28)	&-1.662(11)	&\multicolumn{1}{c}{--}\\
		\multicolumn{2}{l}{Density}	&5.424(33)	&6.845(47)	&-2.114(20)	&\multicolumn{1}{c}{--}\\
		\multicolumn{2}{l}{Temperature}	&6.357(64)	&10.72(14)	&-1.100(20)	&\multicolumn{1}{c}{--}\\
		\hline
		\multirow{1}{*}{Velocity}	&$W''_1$	&3.707(13)	&3.748(16)	&\multirow{2}{*}{0.0990(51)}	&\multirow{2}{*}{0.557(45)}\\
			&$W''_2$	&5.26(13)	&5.42(11)	&	&\\
		\cline{2-6}
		\multicolumn{2}{r}{$W''_\text{II}$}	&4.13(13)\tablefootmark{b}	&4.47(11)\tablefootmark{b}	&\multicolumn{1}{c}{--}	&\multicolumn{1}{c}{--}\\
		\hline
	\end{tabular}
	\tablefoot{
		\tablefoottext{a}{Values in their respective units \si{nT}, \SI{e2}{\km\per\s}, \si{\per\cm\cubed} and \SI{e4}{\K}.}
		\tablefoottext{b}{Velocity median and mean \SI{1}{\au} values for the resulting function. Error estimates derived from the individual fit part errors.}
	}
\end{table*}
% Vdbl fit:
% medians = 370.7, 526
% means = 374.8, 542
% c = 0.557
% via calculate_twolognormal_median.py:
% mean = 446.64
% median = 413.33

@@

The fit models seem to resemble the data quite well (Fig.~\ref{fig:mixed_fit_fixed_4_paper_f_plot}). The magnetic field strength frequency is more focused (around \SI{40}{nT}) at the lower distance boundary than the model's is. This is expected because of our fixed distance independent shape approximation.
\begin{figure*}
	\includegraphics[width=18cm]{figures/mixed_fit_fixed_4_paper_f_plot.pdf}
	\caption{Solar wind parameter's frequency distributions over solar distance. Plotted are the binned Helios data and the exponential lognormal fit model (double lognormal for the velocity) with their median values (white).}
	\label{fig:mixed_fit_fixed_4_paper_f_plot}
\end{figure*}

The velocity and temperature models' upper values have a higher frequency than the data shows. This is due to the systematic fit discrepancy of the lognormal distribution's high value tails (see zoom box in Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}).\\

Possible differing scaling laws at smaller heliocentric distances are discussed in Sect.~\ref{sec:model_extrapolation_to_psp_orbital_time_and_position}.\\

This radial solar wind model represents the Helios time frame around the rise of solar cycle~21.\\

study cycle variation of exponent\\
indications for that: figure! + theory + studies/papers...\\
assuming the distance scaling laws are time independent, which we reason from that; it is an approximation\\

@@

to get the combined model function $W'''(x,ssn,r)$. And for the velocity $W_\text{II}'''(x,ssn,r)$ with the double lognormal function.\\

%model errors
give model errors sizes/uncertainties (sigmas); $W'''_\text{err}(x,ssn,r)$, $x_\text{med\_err}(ssn,r)$\\

model requirements:\\
- modeling the frequency distribution with sufficient accuracy -> error size (changing distribution shape with distance)\\
- modeling the distance dependency with sufficient accuracy -> error size (different scaling law at smaller distances?)\\
- possibility to extrapolate model down to \SI{0.0459}{\au}\\

%model limits
model limits (spherical coordinates):\\
- heliocentric distance range 0.29--0.98~au\\
- rotational symmetry\\
- confined to ecliptic +-7.2 degrees\\
model constrictions:\\
- solar distance dependency function\\
- frequency distribution functions\\
- neglected influence from heliolatitude variation\\

\section{Model extrapolation to PSP orbital time and position}
\label{sec:model_extrapolation_to_psp_orbital_time_and_position}
To estimate the solar wind environment at PSP's planned orbital positions during its mission time, SSN predictions are included into the general solar wind model and extrapolations to the PSP perihelion region are performed.\\

@@

For the prediction of the next solar cycle~25 we simply assume a course similar to the last and thus we shift the last cycle by 11~years. Additionally we consider the two alternatives of half and twice its amplitude. The SSN for PSP's first perihelion will be small---certainly below 20, whereas its nearest perihelia, which commence at the height of cycle~25, will have as of now almost unpredictable SSN amplitudes (see Fig.~\ref{fig:SPP_orbit_predicted_SSN_overview_e_plot})\\
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/SPP_orbit_predicted_SSN_overview_e_plot.pdf}}
	\caption{PSP's solar distance during its mission time (top). Consecutive Venus flybys bring its perihelia nearer to the Sun. Actual and predicted SSN (bottom), i.e. SIDC 13-month smoothed monthly actual SSN, SIDC prediction, SWPC prediction and simply by 11~years shifted SSN from previous cycle~24, together with two alternative trends of half and twice its amplitude.}
	\label{fig:SPP_orbit_predicted_SSN_overview_e_plot}
\end{figure}

[=> Implications for the sw estimation...]\\

>>>>

The fitting of $W_\text{II}(x)$ to the velocity's frequency distribution gives the values of the now five fit parameters ($c$, $x_\text{med,1}$, $x_\text{avg,1}$, $x_\text{med,2}$ and $x_\text{avg,2}$), which are also listed in Table~\ref{tab:lognormal_fit_parameters} together with the median and mean of the composed distribution, which can be derived via solving
\begin{align}
	\int W_\text{II}(x)\,\text{d}x = 0	&	&\text{and}	&	&\int x\,W_\text{II}(x)\,\text{d}x = 0
\end{align}
respectively.\\
As anticipated, this more complex fit function is more accurate in describing the velocity's frequency distribution (see Fig.~\ref{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}).
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/histogram_fits_V_a_zoom_dbl_paper_pdfplot.pdf}}
	\caption{Plot of the velocity's frequency distribution (same as in Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}) and its compositional lognormal fit. The fit's median and mean values and its two fit parts are indicated as well. The inset only has a zoomed-in frequency axis, its x-axis stays the same.}
	\label{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}
\end{figure}
For this reason we keep using the double lognormal ansatz for the velocity frequency fits in the following sections.\\
In this static model the slow and fast part contribute almost equally ($c \approx 0.5$), which of course is only valid for this kind of long-term average. At different times in a solar cycle their contributions vary strongly.

@@

%solar activity = SSN $\neq$ solar cycle\\
The solar wind velocity and its close friends density and temperature are known to depend on the state of the solar cycle \citep{Schwenn1983}, which is why they follow the SSN indirectly (with time lag).
%density modulation with SSN \citep{Schwenn1983} p.~499\\
Thus we derive the correlation coefficients for different time lags between solar wind parameters and SSN (see Fig.~\ref{fig:OMNI_yearly_ssn_correlation_c_plot}).
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/OMNI_yearly_ssn_correlation_c_plot.pdf}}
	\caption{Plot of the solar wind parameter yearly medians from OMNI data and the yearly SSN from the \citet{sidc} with cycle number (top). Their correlation coefficients with the yearly SSN are calculated for time lags back to -15 years (bottom).}
	\label{fig:OMNI_yearly_ssn_correlation_c_plot}
\end{figure}

The highest correlation coefficient for the magnetic field strength is 0.728, which is without lag time and the highest of all solar wind parameters. This is anticipated because the SSN is directly proportional to the magnetic flux \citep{Smith2003}.\\
Velocity and temperature have a lag time of 3~years with their maximal correlation coefficients (0.453 and 0.540). The density has a lag time of 6~years (0.468), which is in agreement with the by \citet{Bougeret1984} documented density anticorrelation with SSN.	%p.~406\\

% maximal correlation coefficients and their lag times:\\
% magnetic field strength: lag 0, 0.728\\
% velocity: lag -3, 0.453\\
% density: lag -6, 0.468\\
% temperature: lag -3, 0.540\\
As expected, the correlation coefficients' amplitudes of all parameters decline with increasing lag time and show a frequency of about 11~years.

@@

Naturally, the fit models match with the general data trends, though single year variations are not able to be replicated by the model (e.g. the high velocity and temperature values in 1974, 1994 and 2003) (see Fig.~\ref{fig:OMNI_yearly_BVdblNTSSN_fit_e_plot}).
\begin{figure*}
	\includegraphics[width=18cm]{figures/OMNI_yearly_BVdblNTSSN_fit_e_plot.pdf}
	\caption{Solar wind parameter data frequencies, lognormal fit models with their median values (white) and the corresponding yearly SSN (grey) over the OMNI time period 1963--2016. The for the models shifted SSN is indicated by a black line. The velocity median is derived from the SSN weighted constant lognormal parts (dotted).}
	\label{fig:OMNI_yearly_BVdblNTSSN_fit_e_plot}
\end{figure*}
The comparison with the yearly data median values over SSN shows that the from the model obtained medians have a quite similar slope (see Fig.~\ref{fig:OMNI_yearly_BVNTvsSSN_a}).
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/OMNI_yearly_BVNTvsSSN_a.pdf}}
	\caption{Solar wind parameter median over lagged SSN. The yearly data medians (+) with their weighted linear fit (solid) are obtained from OMNI data. The error bars denote the SSN standard deviation and the relative weight from the yearly data coverage. The SSN dependent median (\ref{eq:median_with_ssn}) is from the lognormal model fit (dashed). For the velocity the median is derived from the SSN weighting of the slow and fast model parts, whose magnitudes are SSN independent (dotted).}
	\label{fig:OMNI_yearly_BVNTvsSSN_a}
\end{figure}

Again, the velocity gets a special treatment with the double lognormal distribution (\ref{eq:double_lognormal_fit_function}). It is known that slow and fast solar wind stream occurrence rates follow the solar cycle, yet their magnitudes stay fairly stable (cite?). Thus we keep the two velocity components' positions constant and vary instead their balance with the SSN:
\begin{align}
	c(ssn) &= c_a \cdot ssn + c_b\,.	\label{eq:balance_with_ssn}
\end{align}
The fit result (see Table~\ref{tab:ssn_fit_parameters}) is a model in which three years after solar cycle minimum (SSN of zero) the slow solar wind has a share of almost two-thirds and decreases further with increasing SSN (see Fig.~\ref{fig:Vdbl_SSN_ratio_f_plot}).\\
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/Vdbl_SSN_ratio_f_plot.pdf}}
	\caption{Balance of slow to fast solar wind over the by 3~years lagged SSN. The yearly ratios (+) and their weighted linear fit (solid) are obtained from OMNI data with a threshold velocity of $v_\text{th} = \SI{400}{\km\per\s}$. The error bars denote the SSN standard deviation and the relative weight from the yearly data coverage. The model's balance parameter (\ref{eq:balance_with_ssn}) and derived ratio (same threshold) are plotted as dashed and dotted lines.}
	\label{fig:Vdbl_SSN_ratio_f_plot}
\end{figure}
% 3 years after minimum (SSN of zero) the slow wind has its maximal share of 0.638(32). 3 years after maximum (SSN of 200) it has only a share of 0.278(37).\\
To compare the ratios of slow to fast wind between model and data, we simply apply the commonly used constant velocity threshold of $v_\text{th} = \SI{400}{\km\per\s}$ (cite?). The linear fit to the yearly data ratio and the derived model ratio are quite similar (see Fig.~\ref{fig:Vdbl_SSN_ratio_f_plot}). Specific velocity thresholds between slow and fast solar wind cannot be directly compared with the to some degree steeper balance parameter of this model. The model's balance may represent the actual ratios of the solar wind types in a more realistic way than a specific velocity threshold does, since the velocity ranges of both types overlap \citep{McGregor2011b}.
%(i.e. there exists slow wind with properties of fast wind and vice versa)

@@

%Helios data biased towards solar minimum
Using this data, we also have to keep in mind that its time coverage is unequally distributed over the solar cycle. Dividing the data by the transition from cycle minimum to maximum (mid 1977) and considering the data gap distributions, the Helios data covers about \SI{68}{\percent} during cycle minimum whereas during maximum only \SI{38}{\percent}.

@@

For calculating the median and mean values at different solar distances the data is binned into \SI{0.01}{\au} bins, which is also the native precision in this data set.

\subsection{Exponential fitting}
\label{sec:exponential_fitting}
%why exponential function?
An exponential distance behavior is expected from all four parameters (cites?). Therefore we use the exponential function
\begin{align}
	x(r) = d\,r^e	\label{eq:exponential_function}
\end{align}
with the solar distance $r$ for the regression fit of the median and mean. The fits are weighted by data counts per bin.
%fit result table and figure
With $r$ in astronomical units we get the fit coefficients ($d_\text{med}$, $d_\text{avg}$, $e_\text{med}$ and $e_\text{avg}$) as given in Table~\ref{tab:mean_median_fit_parameter}.
\begin{table*}
	\caption{Fit coefficients for the median and mean solar distance dependencies of the four parameters from the combined Helios data set. The errors in brackets are the estimated standard deviations of each fit parameter. The crossing distance is the point where the fitted median and mean intersect. The year variation is the weighted standard deviation from all yearly fitted exponents.}
	\label{tab:mean_median_fit_parameter}
	\centering
	\begin{tabular}{l
	S[table-format = 1.3(2)]
	S[table-format = +1.3(2)]
	@{}c
	S[table-format = 1.3(2)]
	S[table-format = +1.3(2)]
	S[table-format = 1.1(2)e1]
	S[table-format = 1.3]}
		\hline\hline
		\multirow{2}{*}{Parameter}	&\multicolumn{2}{c}{Median}	&	&\multicolumn{2}{c}{Mean}	&\multicolumn{1}{c}{Crossing distance}	&\multicolumn{1}{c}{Year variation}\\
		\cline{2-3}	\cline{5-6}
			&\multicolumn{1}{c}{$d_\text{med}$\tablefootmark{a}}	&\multicolumn{1}{c}{$e_\text{med}$}	&	&\multicolumn{1}{c}{$d_\text{avg}$\tablefootmark{a}}	&\multicolumn{1}{c}{$e_\text{avg}$}	&\multicolumn{1}{c}{[\si{\au}]}	&\multicolumn{1}{c}{$\Delta e$}\\
		\hline
		Magnetic field	&5.377(92)	&-1.655(17)	&	&6.05(10)	&-1.546(18)	&0.339(11)	&0.11\\
		Velocity	&4.107(28)	&0.058(13)	&	&4.356(24)	&0.049(10)	&0.7(83)e3	&0.012\\
		Density		&5.61(27)	&-2.093(46)	&	&7.57(30)	&-2.010(38)	&0.027(73)	&0.072\\
		Temperature	&7.14(23)	&-0.913(39)	&	&9.67(21)	&-0.792(28)	&0.082(85)	&0.005\\
		\hline
	\end{tabular}
	\tablefoot{
		\tablefoottext{a}{Values in their respective units \si{nT}, \SI{e2}{\km\per\s}, \si{\per\cm\cubed} and \SI{e4}{\K}.}
	}
\end{table*}
% source year variation: fit_out_yearly_gradients_table_b_stats.txt

% comparison with other Helios studies (Schwenn, Bougeret)
As expected, the velocity exponents match with those from \citet{Schwenn1983,Schwenn1990}, who derived the distance dependencies for both Helios spacecraft separately ($v_\text{H1}(r) \propto r^{0.083}$ and $v_\text{H2}(r) \propto r^{0.036}$).
%Helios results, radial gradients see \citet{Schwenn1990} p.~155\\
Likewise, the density exponents agree well with the Helios plasma density model, which \citet{Bougeret1984} derived from Helios data and normed to the 1976 \SI{1}{\au} density ($n(r) = 6.14\,r^{-2.10}\,\text{cm}^{-3}$).

@@

As mentioned before, we set the exponents of median and average to be identical. Implementing the exponential distance dependency~(\ref{eq:exponential_function}) into the lognormal function (\ref{eq:single_lognormal_fit_function}), we get three fit parameters ($d'_\text{med}$, $d'_\text{avg}$ and the common exponent $e'$). Naturally, we use the double lognormal function~(\ref{eq:double_lognormal_fit_function}) for the velocity distribution fit, resulting in $W''_\text{II}(x,r)$. The additional fit parameters are the balancing parameter $c'$ and from the second lognormal part $d'_\text{med,2}$ and $d'_\text{avg,2}$. The resulting fit coefficients for the four solar wind parameters are presented in Table~\ref{tab:extrapolation_model_fit_parameters}.
\begin{table*}
	\caption{Fit coefficients from the single lognormal exponential function, respectively double lognormal for the velocity (combined Helios data). The errors in brackets are the estimated standard deviations of each fit parameter.}
	\label{tab:extrapolation_model_fit_parameters}
	\centering
	\sisetup{table-figures-integer=1, table-figures-decimal=3, table-figures-exponent=0}
	\begin{tabular}{l c
	S[table-format = 1.3(2), table-space-text-post = a, table-align-text-post = false]
	S[table-format = 1.3(2), table-space-text-post = a, table-align-text-post = false]
	S[table-format = +1.4(2)]
	S[table-format = 1.3(2)]}
		\hline\hline
		\multicolumn{2}{l}{\multirow{2}{*}{Parameter}}	&\multicolumn{1}{c}{Median\tablefootmark{a}}	&\multicolumn{1}{c}{Mean\tablefootmark{a}}	&\multicolumn{1}{c}{Exponent}	&\multicolumn{1}{c}{Balance}\\
			&	&\multicolumn{1}{c}{$d'_\text{med}$}	&\multicolumn{1}{c}{$d'_\text{avg}$}	&\multicolumn{1}{c}{$e'$}	&$c'$\\
		\hline
		\multicolumn{2}{l}{Magnetic field}	&5.358(25)	&5.705(28)	&-1.662(11)	&\multicolumn{1}{c}{--}\\
		\multicolumn{2}{l}{Density}	&5.424(33)	&6.845(47)	&-2.114(20)	&\multicolumn{1}{c}{--}\\
		\multicolumn{2}{l}{Temperature}	&6.357(64)	&10.72(14)	&-1.100(20)	&\multicolumn{1}{c}{--}\\
		\hline
		\multirow{1}{*}{Velocity}	&$W''_1$	&3.707(13)	&3.748(16)	&\multirow{2}{*}{0.0990(51)}	&\multirow{2}{*}{0.557(45)}\\
			&$W''_2$	&5.26(13)	&5.42(11)	&	&\\
		\cline{2-6}
		\multicolumn{2}{r}{$W''_\text{II}$}	&4.13(13)\tablefootmark{b}	&4.47(11)\tablefootmark{b}	&\multicolumn{1}{c}{--}	&\multicolumn{1}{c}{--}\\
		\hline
	\end{tabular}
	\tablefoot{
		\tablefoottext{a}{Values in their respective units \si{nT}, \SI{e2}{\km\per\s}, \si{\per\cm\cubed} and \SI{e4}{\K}.}
		\tablefoottext{b}{Velocity median and mean \SI{1}{\au} values for the resulting function. Error estimates derived from the individual fit part errors.}
	}
\end{table*}
% Vdbl fit:
% medians = 370.7, 526
% means = 374.8, 542
% c = 0.557
% via calculate_twolognormal_median.py:
% mean = 446.64
% median = 413.33

@@

The fit models seem to resemble the data quite well (Fig.~\ref{fig:mixed_fit_fixed_4_paper_f_plot}). The magnetic field strength frequency is more focused (around \SI{40}{nT}) at the lower distance boundary than the model's is. This is expected because of our fixed distance independent shape approximation.
\begin{figure*}
	\includegraphics[width=18cm]{figures/mixed_fit_fixed_4_paper_f_plot.pdf}
	\caption{Solar wind parameter's frequency distributions over solar distance. Plotted are the binned Helios data and the exponential lognormal fit model (double lognormal for the velocity) with their median values (white).}
	\label{fig:mixed_fit_fixed_4_paper_f_plot}
\end{figure*}
The velocity and temperature models' upper values have a higher frequency than the data shows. This is due to the systematic fit discrepancy of the lognormal distribution's high value tails (see zoom box in Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}).

\subsection{Distance scaling law variations}
\label{sec:distance_scaling_law_variations}

relocate section to 5.3?\\

This radial solar wind model represents the Helios time frame around the rise of solar cycle~21. It is known that the solar wind parameters' magnitudes vary with solar activity. Looking at the variation of the yearly distance scaling laws (Fig.~\ref{fig:yearly_gradients_b}), there is no systematic variation for the magnetic field. The exponents of velocity and temperature seem to follow the SSN and the density not...
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/yearly_gradients_b.pdf}}
	\caption{Helios year variation of the solar wind parameters' fit exponents and SIDC 13-month smoothed monthly SSN. In this plot the year 1974 is omitted because the 21 days covered only }
	\label{fig:yearly_gradients_b}
\end{figure}

For simplicity we assume that the distance scaling laws are time independent and account for this approximation with including the calculated exponents' yearly variations (Table~\ref{tab:mean_median_fit_parameter}) as uncertainities.\\

make remark about influence of unequal solar cycle coverage... (sect.~5.1)\\

Possible differing scaling laws at smaller heliocentric distances are discussed in Sect.~\ref{sec:model_extrapolation_to_psp_orbital_time_and_position}.

@@

to get the combined model function $W'''(x,ssn,r)$. And for the velocity $W_\text{II}'''(x,ssn,r)$ with the double lognormal function (\ref{eq:double_lognormal_fit_function}).\\

%model errors
%give model errors sizes/uncertainties (sigmas); $W'''_\text{err}(x,ssn,r)$, $x_\text{med\_err}(ssn,r)$\\

% model requirements:\\
% - modeling the frequency distribution with sufficient accuracy -> error size (changing distribution shape with distance)\\
% - modeling the distance dependency with sufficient accuracy -> error size (different scaling law at smaller distances?)\\
% - possibility to extrapolate model down to \SI{0.0459}{\au}\\

%model limits
empirical model limits (spherical coordinates):\\
- heliocentric distance range \SIrange{0.29}{0.98}{\au}\\
- rotational symmetry\\
- confined to ecliptic (\SI{+-7.2}{\degree} HGI)\\
model constrictions:\\
- solar distance dependency function\\
- frequency distribution functions\\
- neglected influence from heliolatitude variation\\

\section{Model extrapolation to PSP orbital time and position}
\label{sec:model_extrapolation_to_psp_orbital_time_and_position}
To estimate the solar wind environment at PSP's planned orbital positions during its mission time, SSN predictions are included into the general solar wind model and extrapolations to the PSP perihelion region are performed.

@@

For the prediction of the next solar cycle~25 we simply assume a course similar to the last and thus we shift the last cycle by 11~years. Additionally we consider the two alternatives of half and twice its amplitude. The SSN for PSP's first perihelion will be small---certainly below 20, whereas its nearest perihelia, which commence at the height of cycle~25, will have as of now almost unpredictable SSN amplitudes (see Fig.~\ref{fig:SPP_orbit_predicted_SSN_overview_e_plot}).
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/SPP_orbit_predicted_SSN_overview_e_plot.pdf}}
	\caption{PSP's solar distance during its mission time (top). Consecutive Venus flybys bring its perihelia nearer to the Sun. Actual and predicted SSN (bottom), i.e. SIDC 13-month smoothed monthly actual SSN, SIDC prediction, SWPC prediction and simply by 11~years shifted SSN from previous cycle~24, together with two alternative trends of half and twice its amplitude.}
	\label{fig:SPP_orbit_predicted_SSN_overview_e_plot}
\end{figure}
